---
title: "Direct seeded rice and transplanted rice"
author: "Luis Fernando Delgado"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_download: true
    theme: paper
    df_print: tibble
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
fig.dim = c(7, 4)
library(pacman)
pacman::p_load(tidyverse, DT, agriutilities, emmeans, multcomp, multcompView, 
               factoextra, FactoMineR, readxl, agriutilities)
```

# Loading data Direct seeded rice and transplanted rice

```{r}
transplanted <- read_excel("data/STAT DSR, TPR.xlsx", 
    sheet = "TPR mean")

Direct_seeded <- read_excel("data/STAT DSR, TPR.xlsx", 
    sheet = "DSR mean")
```

# Correlation plot - Transplanted

```{r}
transplanted_cor <- gg_cor(
  colours = c("red", "white", "blue"),
  data = transplanted,
  label_size = 3
  )

print(transplanted_cor)

ggsave(paste0("images\\transplanted_cor", ".png"),
    plot = transplanted_cor, units = "in", dpi = 300, width = 8, height = 6
  )
```

# Correlation plot - Direct seeded

```{r}
Direct_seeded_cor <- gg_cor(
  colours = c("red", "white", "blue"),
  data = Direct_seeded,
  label_size = 3
  )

print(Direct_seeded_cor)

ggsave(paste0("images\\Direct_seeded_cor", ".png"),
    plot = Direct_seeded_cor, units = "in", dpi = 300, width = 8, height = 6
  )
```

# PCA Transplanted - Direct seeded

```{r message=FALSE, warning=FALSE}
pca_transplanted <- transplanted %>% dplyr::select(-Treatment) %>% 
  column_to_rownames("Genotype") %>% PCA(scale.unit = T, graph = F)

pca_direct_seeded <- Direct_seeded %>% dplyr::select(-Treatment) %>% 
  column_to_rownames("Genotype") %>% PCA(scale.unit = T, graph = F)
```

## Correlation circle - Transplanted rice
```{r}
# VAR
var_plot <- fviz_pca_var(pca_transplanted, col.var="contrib", repel = T) +
  labs(title = "PCA transplanted rice")

print(var_plot)

ggsave(paste0("images/PCA_var_transplanted", ".png"), 
       plot = var_plot, units = "in", dpi = 300, width = 8, height = 6)
```

## Correlation circle - Direct seeded rice
```{r}
# VAR
var_plot_directed <- fviz_pca_var(pca_direct_seeded, col.var="contrib", repel = T) +
  labs(title = "PCA direct seeded")

print(var_plot_directed)

ggsave(paste0("images/PCA_var_directed", ".png"), 
       plot = var_plot_directed, units = "in", dpi = 300, width = 8, height = 6)
```



## Individuals Transplanted rice

```{r}
# IND
IND_Transplanted_rice <- fviz_pca_ind(pca_transplanted, repel = T, alpha.ind = 0.3, col.ind = "grey20", labelsize = 2) +
  ggtitle(paste0("Individuals - PCA - Transplanted rice"))

print(IND_Transplanted_rice)

ggsave(paste0("images/PCA_ind_Transplanted_rice", ".png"), 
       plot = IND_Transplanted_rice, units = "in", dpi = 300, width = 8, height = 6)
```

## Individuals Direct seeded rice

```{r}
# IND
IND_Direct_rice <- fviz_pca_ind(pca_direct_seeded, repel = T, alpha.ind = 0.3, col.ind = "grey20", labelsize = 2) +
  ggtitle(paste0("Individuals - PCA - Direct seeded rice"))

print(IND_Direct_rice)

ggsave(paste0("images/PCA_ind_direct_rice", ".png"), 
       plot = IND_Direct_rice, units = "in", dpi = 300, width = 8, height = 6)
```

## Biplot transplated rice

```{r}
BI_transplanted <- fviz_pca_biplot(pca_transplanted, repel = T, alpha.ind = 0.3, 
                                   col.ind = "grey30", labelsize = 3, 
                                   col.var = "black", geom = c("point")) +
  ggtitle(paste0("Biplot - PCA - Transplanted rice")) 


print(BI_transplanted)

ggsave(paste0("images/biplot_transplanted", ".png"), 
       plot = BI_transplanted, units = "in", dpi = 300, width = 8, height = 6)
```

## Biplot direct seeded rice

```{r}
BI_direct <- fviz_pca_biplot(pca_direct_seeded, repel = T, alpha.ind = 0.3, 
                                   col.ind = "grey30", labelsize = 3, 
                                   col.var = "black", geom = c("point")) +
  ggtitle(paste0("Biplot - PCA - Direct seeded rice")) 


print(BI_direct)

ggsave(paste0("images/biplot_direct", ".png"), 
       plot = BI_direct, units = "in", dpi = 300, width = 8, height = 6)
```

# Cluster analysis - Transplanted rice

```{r warning=FALSE, message=FALSE}

#cluster tree
hcpc_transplanted <- HCPC(pca_transplanted, nb.clust = 0, method = "ward", graph = FALSE)

cluster_transplated <- fviz_dend(hcpc_transplanted, 
                                cex = 0.3, 
                                palette = "lancet", 
                                rect = TRUE, 
                                rect_fill = TRUE, 
                                rect_border = "lancet", 
                                labels_track_height = 0.1, 
                                horiz = T) +
  ggtitle("Cluster Transplanted rice")

print(cluster_transplated)

ggsave(paste("images/cluster_transplanted_dend", ".png", sep = "_"), 
       plot = cluster_transplated, units = "in", dpi = 300, width = 8, height = 6)
          
```

## Cluster under PCA - Transplanted rice

```{r message=FALSE, warning=FALSE}
hcpc_pca_transplanted <- fviz_cluster(hcpc_transplanted,
             repel = TRUE,
             show.clust.cent = FALSE,
             labelsize = 4,
             ellipse.type = "convex") +
  ggtitle("Cluster Transplanted rice")

hcpc_pca_transplanted

ggsave(paste("images/cluster_transplanted", ".png", sep = "_"), 
       plot = hcpc_pca_transplanted, units = "in", dpi = 300, width = 8, height = 6)
```

## Significant variables in each cluster

The significant variables are shown below. To understand this table, it is necessary to compare the mean in the category (cluster) to the overall population mean per variable.

```{r}
hcpc_transplanted$desc.var$quanti
```

# Cluster assignment to each genotype - Transplanted rice

```{r}
cluster_assigned <- hcpc_transplanted$data.clust

cluster_assigned <- cluster_assigned %>% rownames_to_column("Genotype") %>% 
  add_column("Treatment" = "transplanted") %>% 
  dplyr::select(Genotype, Treatment, clust, everything()) %>% 
  mutate(across(where(is.numeric), ~ round(., 2)))

cluster_assigned %>% datatable(extensions = 'Buttons',
            options = list(dom = 'Blfrtip', pageLength = 6, autoWidth = TRUE,
                           buttons = c('copy', 'csv', 'excel'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))

```


# Cluster analysis - Direct seeded rice

```{r warning=FALSE, message=FALSE}

#cluster tree
hcpc_direct <- HCPC(pca_direct_seeded, nb.clust = 0, method = "ward", graph = FALSE)

cluster_direct <- fviz_dend(hcpc_direct, 
                                cex = 0.3, 
                                palette = "lancet", 
                                rect = TRUE, 
                                rect_fill = TRUE, 
                                rect_border = "lancet", 
                                labels_track_height = 0.1, 
                                horiz = T) +
  ggtitle("Cluster Direct seeded rice")

print(cluster_direct)

ggsave(paste("images/cluster_Direct_dend", ".png", sep = "_"), 
       plot = cluster_direct, units = "in", dpi = 300, width = 8, height = 6)
          
```

## Cluster under PCA - Transplanted rice

```{r message=FALSE, warning=FALSE}
hcpc_pca_direct <- fviz_cluster(hcpc_direct,
             repel = TRUE,
             show.clust.cent = FALSE,
             labelsize = 4,
             ellipse.type = "convex") +
  ggtitle("Cluster Direct seeded rice")

hcpc_pca_direct

ggsave(paste("images/cluster_direct", ".png", sep = "_"), 
       plot = hcpc_pca_direct, units = "in", dpi = 300, width = 8, height = 6)
```

## Significant variables in each cluster - Direct seeded rice

The significant variables are shown below. To understand this table, it is necessary to compare the mean in the category (cluster) to the overall population mean per variable. Direct seeded rice population formed 3 clusters

```{r}
hcpc_direct$desc.var$quanti
```

# Cluster assignment to each genotype - Transplanted rice

```{r}
cluster_assigned_diret <- hcpc_direct$data.clust

cluster_assigned_diret <- cluster_assigned_diret %>% rownames_to_column("Genotype") %>% 
  add_column("Treatment" = "Direct_seeded") %>% 
  dplyr::select(Genotype, Treatment, clust, everything()) %>% 
  mutate(across(where(is.numeric), ~ round(., 2)))

cluster_assigned_diret %>% datatable(extensions = 'Buttons',
            options = list(dom = 'Blfrtip', pageLength = 6, autoWidth = TRUE,
                           buttons = c('copy', 'csv', 'excel'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))



```

# Pull together both data sets

```{r}
both <- transplanted %>% bind_rows(Direct_seeded)

pca_both <- both %>% dplyr::select(-c(Genotype)) %>% PCA(scale.unit = T, 
                                                     graph = F, quali.sup = 1)
```

## Circle correlations both treatments
```{r}

both_treatment <- fviz_pca_ind(pca_both, label="none", habillage = 1, addEllipses=TRUE)

print(both_treatment)

ggsave(paste("images/ind_both_treatment", ".png", sep = "_"), 
       plot = both_treatment, units = "in", dpi = 300, width = 8, height = 6)
```

## Biplot both treatments
```{r}
BI_both <- fviz_pca_biplot(pca_both, label="var", habillage=1,
               addEllipses=TRUE, ellipse.level=0.95)

print(BI_both)

ggsave(paste("images/BI_both_treatment", ".png", sep = "_"), 
       plot = BI_both, units = "in", dpi = 300, width = 8, height = 6)


```



